---
title: "Untitled"
author: "Elyza Pilipaityte"
date: "2024-07-08"
output: html_document
---

```{r}
rm(list=ls())
```

Libraries
```{r}
library(tidyverse)
library(lme4)
library(sjPlot)
library(rfishbase)
library(car)

My_theme <- theme(
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA, linewidth = 1.25),
  strip.background = element_rect(fill = "white", color = "white", linewidth = 1.25),
  text = element_text(size = 16, family = "gillsans"),
  panel.grid.major = element_blank(),  # Remove major grid lines
  panel.grid.minor = element_blank()   # Remove minor grid lines
)
```

popgrowth tables

```{r}
# getting species names from fishbase
fish <- species_names()
species_list <- fish$Species
# extracting available growth data for all species
spp_data <- popgrowth(species_list)
# number of species
length(unique(spp_data$Species))

# selecting only necessary variables and excluding doubtful estimates
data <- spp_data %>% select(Species, method = Data, TLinfinity, K, Auxim, Sex, Type, Loo, to) %>% filter(Auxim != "doubtful")

# any missing aging method
colSums(is.na(data))
# removing missing aging method
data <- data %>% drop_na(method)

# aging methods
table(data$method)
# renaming aging methods
data$method <- ifelse(data$method %in% c("annuli on few otoliths", "annuli on many otoliths", "annuli on otoliths", "daily otolith rings", "otoliths"), "otoliths",
ifelse(data$method %in% c("others", "several data types", "other annual rings", "various data types"), "others",
ifelse(data$method %in% c("scale annual rings", "scales", "scale"), "scales",
ifelse(data$method %in% c("Lmax Lm tm", "Lmax tmax", "length-frequencies"), "growth",
ifelse(data$method %in% c("direct observation", "tagging/recapture"), "direct_observ",
data$method)))))

# filtering only scale or otolith method
data <- data %>% filter(method == "scales" | method == "otoliths")
length(unique(data$Species))

# Filter species that have both "scales" and "otoliths" aging methods
species_with_both_methods <- data %>%
  group_by(Species) %>%
  summarise(has_scales = "scales" %in% method,
            has_otoliths = "otoliths" %in% method) %>%
  filter(has_scales & has_otoliths) %>%
  pull(Species)

# Filter the original data to include only species with both methods and exclude doubtful estimations of Linf
data_filtered <- data %>% filter(Species %in% species_with_both_methods) %>% filter(method == "otoliths" | method == "scales")

```

Table to see how many observations in each method for different species

```{r}
# Group by species and method, then count the number of observations
method_counts <- data_filtered %>%
  group_by(Species, method) %>%
  summarise(count = n(), .groups = 'drop')

# Create a table with species names as rows and methods as columns
method_counts_table <- method_counts %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)
```

Many species have only one estimation for one of the methods, I filter the species which have more than 4 estimates using each method.

```{r}
# Filter species that have more than 4 Linf counts using both methods
filtered_species <- data_filtered %>%
  group_by(Species, method) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0) %>%
  filter(scales > 4 & otoliths > 4) %>%
  pull(Species)

# Further filter the original data to include only species meeting the criteria
data_filtered_final <- data_filtered %>%
  filter(Species %in% filtered_species)

table(data_filtered_final$Species)
```

year included in the csv file from FishBase references manually

```{r}
data <- read_csv(file = "rfish_Linf_year.csv")
```

```{r}
str(data)

ggplot(data, aes(x = year, y = method, fill = method)) +
  geom_boxplot() +
  xlab("year") +
  ylab("method") +
  My_theme


data <- data %>% filter(year > 1925)

ggplot(data, aes(x = year, y = method, fill = method)) +
  geom_boxplot() +
  xlab("year") +
  ylab("method") +
  My_theme

ggplot(data, aes(x = year, fill = method)) +
  geom_bar(position = "dodge") +
  xlab("Year") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 10)) +
 My_theme


data_summary <- data %>%
  group_by(year, method) %>%
  summarise(Frequency = n())

ggplot(data_summary, aes(x = year, y = Frequency, color = method)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_smooth(method = "lm", se = FALSE, size = 1.5) +
  xlab("Year") + ylab("Frequency of Linf Observations") +
  My_theme + 
geom_smooth(method = "lm", se = TRUE, size = 1.5)

table(data$year)
table(data$Locality)
table(data$Sex)
table(data$Species)

ggplot(data, aes(x = as.factor(year), y = Loo, color = method)) +
  geom_jitter(size = 2) +
  My_theme +
  facet_wrap(~ Species) +
  scale_x_discrete(breaks = unique(data$year)[seq(1, length(unique(data$year)), by = 10)]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data, aes(x = as.factor(year), y = Loo, color = method, shape = Type)) +
  geom_jitter(size = 2) +
  My_theme +
  facet_wrap(~ Species) +
  scale_x_discrete(breaks = unique(data$year)[seq(1, length(unique(data$year)), by = 10)]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(shape = "Length Type")

ggplot(data, aes(x = Species, y = Loo, fill = method)) +
  geom_boxplot(alpha = 0.6) +
  theme_minimal() +
  xlab("Species") +
  ylab(expression(L[infinity]~"(cm)")) +
  scale_fill_manual(values = c("scales" = "#D55E00", "otoliths" = "#332288")) +
  theme(panel.background = element_rect(fill = "white"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 16), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 55, hjust = 1),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```


Data exploration 

1. Outliers in response and independent variables

```{r}
ggplot(data = data) + 
  geom_point(mapping = aes(x = Loo, y = Species, color=as.factor(method)), position = "jitter", size = 2.5)
# no outliers
```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plot

hist(data$Loo,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")

```

3. Balance of categorical variables

```{r}
table(data$Species)
table(data$method)
table(data$year)
```

4. Multicollinearity among covariates

```{r}
#Calculate Variance Inflation Factor (VIF)
round(vif(lm(Loo ~ method + Species + year,
                     data = data)),2)

round(vif(lm(K ~ method + Species + year,
                     data = data)),2)
```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(data, aes(x = year, y = (Loo), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)") +
  facet_wrap(~method)

ggplot(data, aes(x = year, y = (Linf_stand), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)") +
  facet_wrap(~method)



ggplot(data, aes(x = method, y = (Loo), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)")

ggplot(data, aes(x = method, y = (Linf_stand), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)")
```


Models

```{r}
m0 <- lm(Loo ~ method, data = data)
m1 <- lmer(Loo ~ method + (1|Species), data = data, REML = F)
m2 <- lmer(Loo ~ method + (1|Species) + (1|year), data = data, REML = F)
anova(m1, m2)

m3 <- lmer(Loo ~ method + (1|Species) + (1|year) + (1|Locality), data = data, REML = F)
anova(m2, m3)

m4 <- lmer(Linf_stand ~ method + (1|Species), data = data, REML = F)
m5 <- lmer(Linf_stand ~ method + year + (1|Species), data = data, REML = F)
anova(m4, m5)

m6 <- lmer(Linf_stand ~ method + (1|Species) + (1|year), data = data, REML = F)
m7 <- lmer(Linf_stand ~ method + (1|Species) + (1|year) + (1|Locality), data = data, REML = F)
anova(m4, m6)
anova(m7, m6)

anova(m2, m6)
```

```{r}
# Summarise neatly in a table
tab_model(m1,
          show.zeroinf = F,
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)

plot_model(m2)
```

Residuals

```{r}
# 1. Linearity & Homoscedasticity: Residuals vs Fitted plot
plot(m1, which = 1)

# 2. Normality of residuals: Q-Q plot of residuals
qqnorm(residuals(m1))
qqline(residuals(m1))

# 3. Normality of random effects: Q-Q plot for random effects
qqnorm(ranef(m1)$Species[,1])
qqline(ranef(m1)$Species[,1])

# 4. Variance of random effects (Species)
summary(m1)$varcor

hist((residuals(m1)), breaks = 30, main = "Histogram of Residuals")
```

K Model

```{r}
m0 <- lm(K ~ method, data = data)
m1 <- lmer(K ~ method + (1|Species), data = data, REML = F)
m2 <- lmer(K ~ method + (1|Species) + (1|year), data = data, REML = F)
anova(m1, m2)

```


```{r}
# Summarise neatly in a table
tab_model(m1,
          show.zeroinf = F,
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)

```

Residuals

```{r}
# 1. Linearity & Homoscedasticity: Residuals vs Fitted plot
plot(m1, which = 1)

# 2. Normality of residuals: Q-Q plot of residuals
qqnorm(residuals(m1))
qqline(residuals(m1))

# 3. Normality of random effects: Q-Q plot for random effects
qqnorm(ranef(m1)$Species[,1])
qqline(ranef(m1)$Species[,1])

# 4. Variance of random effects (Species)
summary(m1)$varcor

hist((residuals(m1)), breaks = 30, main = "Histogram of Residuals")
```


### REVISION

```{r}



```

