---
title: "Untitled"
author: "Elyza Pilipaityte"
date: "2024-07-08"
output: html_document
---

```{r}
rm(list=ls())
```

Libraries
```{r}
library(tidyverse)
library(lme4)
library(sjPlot)
library(rfishbase)
library(car)

My_theme <- theme(
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA, linewidth = 1.25),
  strip.background = element_rect(fill = "white", color = "white", linewidth = 1.25),
  text = element_text(size = 16, family = "gillsans"),
  panel.grid.major = element_blank(),  # Remove major grid lines
  panel.grid.minor = element_blank()   # Remove minor grid lines
)
```

popgrowth tables

```{r}
# getting species names from fishbase
fish <- species_names()
species_list <- fish$Species
# extracting available growth data for all species
spp_data <- popgrowth(species_list)
# number of species
length(unique(spp_data$Species))

# selecting only necessary variables and excluding doubtful estimates
data <- spp_data %>% select(Species, method = Data, TLinfinity, K, Auxim, Sex, Type, Loo, to) %>% filter(Auxim != "doubtful")

# any missing aging method
colSums(is.na(data))
# removing missing aging method
data <- data %>% drop_na(method)

# aging methods
table(data$method)
# renaming aging methods
data$method <- ifelse(data$method %in% c("annuli on few otoliths", "annuli on many otoliths", "annuli on otoliths", "daily otolith rings", "otoliths"), "otoliths",
ifelse(data$method %in% c("others", "several data types", "other annual rings", "various data types"), "others",
ifelse(data$method %in% c("scale annual rings", "scales", "scale"), "scales",
ifelse(data$method %in% c("Lmax Lm tm", "Lmax tmax", "length-frequencies"), "growth",
ifelse(data$method %in% c("direct observation", "tagging/recapture"), "direct_observ",
data$method)))))

# filtering only scale or otolith method
data <- data %>% filter(method == "scales" | method == "otoliths")
length(unique(data$Species))

# Filter species that have both "scales" and "otoliths" aging methods
species_with_both_methods <- data %>%
  group_by(Species) %>%
  summarise(has_scales = "scales" %in% method,
            has_otoliths = "otoliths" %in% method) %>%
  filter(has_scales & has_otoliths) %>%
  pull(Species)

# Filter the original data to include only species with both methods and exclude doubtful estimations of Linf
data_filtered <- data %>% filter(Species %in% species_with_both_methods) %>% filter(method == "otoliths" | method == "scales")

```

Table to see how many observations in each method for different species

```{r}
# Group by species and method, then count the number of observations
method_counts <- data_filtered %>%
  group_by(Species, method) %>%
  summarise(count = n(), .groups = 'drop')

# Create a table with species names as rows and methods as columns
method_counts_table <- method_counts %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)
```

Many species have only one estimation for one of the methods, I filter the species which have more than 4 estimates using each method.

```{r}
# Filter species that have more than 4 Linf counts using both methods
filtered_species <- data_filtered %>%
  group_by(Species, method) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0) %>%
  filter(scales > 4 & otoliths > 4) %>%
  pull(Species)

# Further filter the original data to include only species meeting the criteria
data_filtered_final <- data_filtered %>%
  filter(Species %in% filtered_species)

table(data_filtered_final$Species)
```

year included in the csv file from FishBase references manually

```{r}
data <- read_csv(file = "rfish_Linf_year.csv")
```

```{r}
str(data)

ggplot(data, aes(x = year, y = method, fill = method)) +
  geom_boxplot() +
  xlab("year") +
  ylab("method") +
  My_theme


data <- data %>% filter(year > 1925)

ggplot(data, aes(x = year, y = method, fill = method)) +
  geom_boxplot() +
  xlab("year") +
  ylab("method") +
  My_theme

ggplot(data, aes(x = year, fill = method)) +
  geom_bar(position = "dodge") +
  xlab("Year") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 10)) +
 My_theme


data_summary <- data %>%
  group_by(year, method) %>%
  summarise(Frequency = n())

ggplot(data_summary, aes(x = year, y = Frequency, color = method)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_smooth(method = "lm", se = FALSE, size = 1.5) +
  xlab("Year") + ylab("Frequency of Linf Observations") +
  My_theme + 
geom_smooth(method = "lm", se = TRUE, size = 1.5)

table(data$year)
table(data$Locality)
table(data$Sex)
table(data$Species)
table(data$method)
table(data$Species, data$method)


ggplot(data, aes(x = as.factor(year), y = Loo, color = method)) +
  geom_jitter(size = 2) +
  My_theme +
  facet_wrap(~ Species) +
  scale_x_discrete(breaks = unique(data$year)[seq(1, length(unique(data$year)), by = 10)]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data, aes(x = as.factor(year), y = Loo, color = method, shape = Type)) +
  geom_jitter(size = 2) +
  My_theme +
  facet_wrap(~ Species) +
  scale_x_discrete(breaks = unique(data$year)[seq(1, length(unique(data$year)), by = 10)]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(shape = "Length Type")

ggplot(data, aes(x = Species, y = Loo, fill = method)) +
  geom_boxplot(alpha = 0.6) +
  theme_minimal() +
  xlab("Species") +
  ylab(expression(L[infinity]~"(cm)")) +
  scale_fill_manual(values = c("scales" = "#D55E00", "otoliths" = "#332288")) +
  theme(panel.background = element_rect(fill = "white"), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 16), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 55, hjust = 1),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```


Data exploration 

1. Outliers in response and independent variables

```{r}
ggplot(data = data) + 
  geom_point(mapping = aes(x = Loo, y = Species, color=as.factor(method)), position = "jitter", size = 2.5)
# no outliers
```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plot

hist(data$Loo,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")

```

3. Balance of categorical variables

```{r}
table(data$Species)
table(data$method)
table(data$year)
```

4. Multicollinearity among covariates

```{r}
#Calculate Variance Inflation Factor (VIF)
round(vif(lm(Loo ~ method + Species + year,
                     data = data)),2)

round(vif(lm(K ~ method + Species + year,
                     data = data)),2)
```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(data, aes(x = year, y = (Loo), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)") +
  facet_wrap(~method)

ggplot(data, aes(x = year, y = (Linf_stand), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)") +
  facet_wrap(~method)



ggplot(data, aes(x = method, y = (Loo), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)")

ggplot(data, aes(x = method, y = (Linf_stand), color = Species)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("method") + ylab("Linf (cm)")
```


Models

```{r}
m0 <- lm(Loo ~ method, data = data)
m1 <- lmer(Loo ~ method + (1|Species), data = data, REML = F)
m2 <- lmer(Loo ~ method + (1|Species) + (1|year), data = data, REML = F)
anova(m1, m2)

m3 <- lmer(Loo ~ method + (1|Species) + (1|year) + (1|Locality), data = data, REML = F)
anova(m2, m3)

m4 <- lmer(Linf_stand ~ method + (1|Species), data = data, REML = F)
m5 <- lmer(Linf_stand ~ method + year + (1|Species), data = data, REML = F)
anova(m4, m5)

m6 <- lmer(Linf_stand ~ method + (1|Species) + (1|year), data = data, REML = F)
m7 <- lmer(Linf_stand ~ method + (1|Species) + (1|year) + (1|Locality), data = data, REML = F)
anova(m4, m6)
anova(m7, m6)

anova(m2, m6)
```

Z transformation
```{r}
data$Linf_z <- scale(data$Loo)
data$K_z <- scale(data$K)

m1 <- lmer(Linf_z ~ method + (1|Species) + (1|K_z), data = data, REML = F)
```

```{r}
# Summarise neatly in a table
tab_model(m1,
          show.zeroinf = F,
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)

plot_model(m2)
```

Residuals

```{r}
# 1. Linearity & Homoscedasticity: Residuals vs Fitted plot
plot(m1, which = 1)

# 2. Normality of residuals: Q-Q plot of residuals
qqnorm(residuals(m1))
qqline(residuals(m1))

# 3. Normality of random effects: Q-Q plot for random effects
qqnorm(ranef(m1)$Species[,1])
qqline(ranef(m1)$Species[,1])

# 4. Variance of random effects (Species)
summary(m1)$varcor

hist((residuals(m1)), breaks = 30, main = "Histogram of Residuals")
```

K Model

```{r}
m0 <- lm(K ~ method, data = data)
m1 <- lmer(K ~ method + (1|Species), data = data, REML = F)
m2 <- lmer(K ~ method + (1|Species) + (1|year), data = data, REML = F)
anova(m1, m2)

```


```{r}
# Summarise neatly in a table
tab_model(m1,
          show.zeroinf = F,
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)

```

Residuals

```{r}
# 1. Linearity & Homoscedasticity: Residuals vs Fitted plot
plot(m1, which = 1)

# 2. Normality of residuals: Q-Q plot of residuals
qqnorm(residuals(m1))
qqline(residuals(m1))

# 3. Normality of random effects: Q-Q plot for random effects
qqnorm(ranef(m1)$Species[,1])
qqline(ranef(m1)$Species[,1])

# 4. Variance of random effects (Species)
summary(m1)$varcor

hist((residuals(m1)), breaks = 30, main = "Histogram of Residuals")
```


### REVISION

```{r}
library(brms)

# Define two formulas for the responses
f1 <- bf(Linf_z ~ method + (1 | Species))
f2 <- bf(K_z ~ method + (1 | Species))

# Fit multi-response model
fit_multi <- brm(f1 + f2 + set_rescor(FALSE),  # set_rescor = FALSE to assume independent residuals
                 data = data,
                 chains = 4, cores = 4,
                 iter = 4000,
                 control = list(adapt_delta = 0.95))


summary(fit_multi)

# Get marginal effects (fixed effects only)
plot(marginal_effects(fit_multi), ask = FALSE)
```

```{r}
# For Loo (L∞)
new_data_Loo <- data.frame(method = c("otoliths", "scales"))
pred_Loo <- as.data.frame(fitted(fit, newdata = new_data_Loo, re_formula = NA, resp = "Loo", summary = TRUE))
pred_Loo$response <- "Loo"
pred_Loo$method <- new_data_Loo$method

# For K
new_data_K <- data.frame(method = c("otoliths", "scales"))
pred_K <- as.data.frame(fitted(fit, newdata = new_data_K, re_formula = NA, resp = "K", summary = TRUE))
pred_K$response <- "K"
pred_K$method <- new_data_K$method

# Combine
pred_df_long <- rbind(pred_Loo, pred_K)


library(ggplot2)
library(tidyr)
library(dplyr)

# Long format for original data
data_long <- data %>%
  pivot_longer(cols = c(Loo, K), names_to = "response", values_to = "value")

# Plot
ggplot(data_long, aes(x = method, y = value)) +
  geom_jitter(width = 0.2, alpha = 0.4, color = "gray") +
  geom_pointrange(data = pred_df_long,
                  aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
                  color = "blue", size = 1.2) +
  facet_wrap(~ response, scales = "free_y") +
  theme_minimal() +
  labs(x = "Ageing Method", y = "Value", title = "Effect of Ageing Method on L∞ and K")

ggplot(pred_df_long, aes(x = method, y = Estimate, color = method)) +
  geom_pointrange(aes(ymin = Q2.5, ymax = Q97.5), size = 1.2) +
  facet_wrap(~ response, scales = "free_y") +
  theme_minimal() +
  labs(title = "Model Estimates of L∞ and K by Ageing Method",
       y = "Estimated Value", x = "Ageing Method") +
  theme(legend.position = "none")


```

```{r}
library(brms)
library(bayesplot)

# Extract posterior samples
posterior <- as_draws_df(fit_multi)

# Select only fixed effects
posterior_selected <- posterior %>%
  dplyr::select(b_Linfz_methodscales, b_Kz_methodscales)

# Plot posterior distributions with 80% and 95% credible intervals
bayesplot::mcmc_areas(
  posterior_selected,
  prob = 0.95,        # 95% CI
  prob_outer = 1.0,   # Full range
  point_est = "mean"
) +
  ggplot2::labs(
    title = "Posterior Distributions of Ageing Method Effect",
    x = "Estimate (z-scale)",
    y = NULL
  )

```

```{r}
# Get conditional effects
ce <- conditional_effects(fit_multi)

# Plot both responses
plot(ce, points = TRUE)[[1]]  # for Linf_z
plot(ce, points = TRUE)[[2]]  # for K_z

```

```{r}
posterior_long <- posterior %>%
  dplyr::select(
    L_inf = b_Linfz_methodscales,
    K = b_Kz_methodscales
  ) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "Parameter", values_to = "Estimate")

ggplot(posterior_long, aes(x = Estimate, fill = Parameter)) +
  geom_density(alpha = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ Parameter, scales = "free") +
  theme_minimal() +
  labs(title = "Posterior Distributions: Effect of Scale Method on Growth Parameters",
       x = "Posterior Estimate", y = "Density")

```



